Mergers in the COSMOS Field

### 2023-11-17 paper-source/main.tex
Major update to code and paper. Code, reclassified stages. Paper, methods section completed and prepping to write up results.

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
% \usepackage{enumitem}

@@

\section{Introduction}\label{introduction}
 Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Specifically, in the theory of $\Lambda$CDM cosmology, where structure is expected to form hierarchically \cite{normal cosmology papers}. These interactions lead to disturbed morphologies \citep{normal disturbed morphology papers}, intense starbursts \citep{normal starburst papers} and, potentially, the rapid quenching of those systems \citep{normal quenching papers}. These process are found to occur not just in our simulations of galaxy interaction \citep{new simulation papers} but across a wide range of observations of interacting systems as well \citep{new observation papers}. There is also evidence to suggest that interaction plays a role in the activation of active galactic nuclei (AGN) \citep{normal papers on AGN activation}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution.

 To completely understand galaxy interaction, and put reliable statistical constraints on its effect on galaxies, we require large samples of interacting galaxies to compare to non-interacting control galaxies. These large samples are surprisingly difficult to come by; with the first interacting galaxy catalogue (the \citet{1966ApJS...14....1A} catalogue) containing only 318 systems. More recent studies have created their own samples, and used them to draw conclusions on the effects of interacting and merging galaxies on galaxy formation. \textbf{Give examples here of such studies}.

 However, we elect to use the newly created interacting galaxy catalogue created by \citet{2023ApJ...948...40O} (hereafter, OR23) which found and published a sample of 21,926 interacting galaxies. This sample was created using the new ESA Datalabs\footnote{ESA Datalabs: link} platform and the new machine learning algorithm \texttt{Zoobot} \citep{2022MNRAS.509.3966W, 2023JOSS....8.5312W} classifying all sources in the Hubble Source Catalogue from the Hubble Space Telescope (HST) archives as to whether they were interacting or not. A rigorous process was then conducted to select likely interacting and merging candidates. They presented this catalogue with initial explorations of the catalogue (such as its limitations) but mainly focused on the methods used to create it.

 In this work, we utilise the new OR23 catalogue and match it with the Cosmic Evolutionary Survey (COSMOS) catalogues and explore the parameters of these systems in a statistically meaningful way. We elect to use COSMOS as it has much flexibility in the parameters we can extract, having many editions of its catalogues using a variety of software to give estimates of stellar masses, star formation rates, luminosity's, and redshift information. For our aims, we use a combination of the COSMOS2015 \citep{2016ApJS..224...24L} and COSMOS2020 \citep{2022ApJS..258...11W} catalogues. The COSMOS2015 catalogue contains excellent estimates of stellar mass and photometric redshift, while the COSMOS2020 catalogue contains information to calculate the star formation rates of each galaxy from estimate luminosity's as well as X-ray flux information to define an AGN.

 Interestingly, there has also been little work done to investigate mergers and interacting galaxies that exist in the COSMOS field, with the most recent work \citet{2022ApJ...940....4S} investigating on 2351 systems throughout the field, focusing on the relationship of projected separation of the galaxy pairs with star formation; finding a significant enhancement in each interactor the smaller the projected separation between them was. Using OR23, we find a sample of 3096 interacting systems that are also in the COSMOS field. We will use these to investigate not only the star formation rate difference, but also the rate of occurance of AGN and the evolution of the nonparametric CAS, Gini and M$_{20}$ parameters.

 Upon matching the OR23 catalogue to COSMOS, we also break the matched systems down into different stages in the interaction, attempting to investigate the changes in star formation rates, AGN activity and morphology as an interaction continues. We create a control sample, and aim to quantify the increase in star formation experienced by an interacting galaxy and also aim to see what this increase is at its height. We also compare the difference in occurrence in AGN between the interacting and control sample, and also speculate on the potential power of these AGN in the comparison. Finally, we measure the change in the regular morphology parameters from \citet{2004AJ....128..163L} throughout the different stages.

 This paper is laid out as follows in Section \ref{data} we fully explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching. Section \ref{staging} describes the methods by which we split the interacting systems in their different stages, calculate the star formation rates of each system as well as describe how we classify a galaxy as an AGN. In Section \ref{results} we full describe and discuss our results, and their implications, before we make concluding remarks in Section \ref{conclusion}.

@@

\section{Data \& Catalogue Matching} \label{data}
\noindent In this work, we utilise the interacting galaxy catalogue created in \citet{2023ApJ...948...40O}. While the creation of this catalogue is fully detailed there, we will briefly describe it here. This catalogue of 21,926 interacting galaxies was created using a Convolutional Neural Network - \textit{Zoobot} - to search all final product data in the \textit{Hubble} Legacy Archive. This was done using single-band \textit{Hubble} images, specificially the \textit{F814W} filter. This choice was motivated by it having the highest number of final product FITS files associated with it. Using the \textit{Hubble} Source Catalogue to extract 92 million sources, these were classified into non-interacting and interacting galaxies. Due to existing issues with using CNNs (see \citet{papers on problems with CNN}) an extremely conservative prediction score from \textit{Zoobot} was required to define an interacting galaxy. This resulted in a very pure, but highly incomplete catalogue of interacting galaxies. Still, this is the largest catalogue of interacting galaxies directly found to date.

However, in the OR23 catalogue, only the coordinates and the \textit{Hubble} Source catalogue are provided. For ancillary data, this catalogue must be matched to different data repositories. This is not a trivial task, as a large proportion of the OR23 catalogue is known to not exist in any existing astrophysical data repository. However, in this work, we opt to match the OR23 catalogue to the Cosmic Evolution Survey (COSMOS) 2020 catalogue \citep{2022ApJS..258...11W}. This contains estimates of stellar masses, photometric redshifts, star formation rates and many line fluxes estimated from LaPhare \citep{LaPhare paper} and EAZY \citep{EAZY paper}. It also contains the observed fluxes for multiple lines from UVISTA and Subaru HyperSuprime-Cam (HSC). For all further measurements of absolute magnitude, we opt to use the HSC infrared band, centered on 7740.58\AA. This is selected as it is the most relevant in detecting radiation from star forming regions over our redshift range.

The two catalogues were matched by searching a 10" radius about each interacting galaxy in OR23. The galaxy closest to the OR23 coordinates in projected separation was defined as the primary galaxy. To find the secondary galaxy, another search radius of 30" was defined about the found primary galaxy coordinates in the COSMOS catalogue. A range of $\pm0.001$ about the primary redshift was taken. Note, this was based on the estimated photometric redshift from EAZY. Once this cut had been applied, the galaxy closest to the primary galaxy in projected separation was taken to be the secondary galaxy. With these criteria, a matched sample of 3,568 primary galaxies were found. For each primary, a secondary was also successfully recovered giving us a total of 7,136 interacting galaxies.

However, duplication existed in this matched catalogue. This was when a primary galaxy in OR23 had been assigned to multiple COSMOS sources or vice versa. This was also true of secondaries, where the secondary galaxy source could be assigned to different primaries. Therefore, a de-duplication was conducted on the COSMOS Identification numbers. The same ID for the primary or secondary could not appear multiple times, as well as no primary or secondary galaxy IDs could match. Upon this de-duplication, we were left with a sample of 3,437 systems or 6,874 interacting galaxies.

\section{Finding Secondaries \& Staging Interacting Galaxies}\label{staging}
In this section, we describe our methodology of finding the secondaries about each galaxy and detail how we classified each interaction into separate stages. The \citet{2023ApJ...948...40O} catalogue only made predictions on interacting systems, not individual galaxies. This was due to the high number of duplication found in the source ids in the HSC. Therefore, very often, any neighbouring secondary galaxy that would have appeared in the catalogue has been merged under a single SourceID related to the primary galaxy. We use the photometric redshifts available in the COSMOS2020 catalogue to search around each primary galaxy for their related secondary. Each interacting pair is then split into four different stages of interaction. This was done by four of the authors of this work. Here, we detail how we maintain statistical robustness and quantify our biases even with multiple classifiers. We also detail the number of galaxies in each stage bin.

\subsection{Secondary Identification}
\noindent As stated previously, the \citet{2023ApJ...948...40O} catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. We achieve this by applying three steps to each system. First, we identify all sources within a 10" radius of our identified primary interacting galaxy. We then create a slice in redshift to find galaxies which are actually nearby in 3D. This is done using a similar parameter as that defined by \citet{Baldry et al. (2006)} when calculating the environment about each galaxy. We define the slice in redshift by $\pm\Delta$zc $= \pm1000$km/s.

Upon applying these two criteria, we are often only left with very few potential candidates to be the secondary interacting galaxy. At this stage, we then order all of our candidates by the measured 3D distance from our primary galaxy using the on-sky separation and their respective redshifts. The closest candidate is then used as the secondary galaxy. To test this method is robust, we conduct two final checks on our candidate interacting systems. First, for all systems we can, we look up the spectroscopic information of each. We find no serious change in the redshift between the photometric and spectroscopic redshifts. We also finally conduct a visual inspection of each candidate interacting galaxy system. We confirm that the potential secondary does exhibit morphological disturbance or is clearly involved in the interaction with the primary.

In a few cases (N), we find that our secondary identification has failed. This is purely due to the secondary already merging with the primary galaxy, and therefore, not being identified as a second source or object in the COSMOS2020 catalogue. Later on, these will be classified as our stage 4 galaxies. We keep these galaxies in our sample, as they are clearly merging, merged or post-merger candidates, however do not attempt to find a secondary for them. Thus, our sample of 3,784 primary interacting galaxies is increased, with their secondaries, to N galaxies.

\subsection{Classifying Stage of Interaction}
The primary purpose of this project is to compare different stages of interaction and investigate changes in their underlying photometric parameterisations. This will give us insights into the interplay between interaction and these underlying processes. Therefore, in this section we describe how we break our interacting galaxies into different stages.

Our sample overlapping with COSMOS20202 is broken down into four different stages. We visually classify each system accounting for the pair separation and distortion. The four different stages are as follows:

% \vspace{-2mm}

\begin{itemize}
    \item Stage 1: Well-separated with no morphological disturbance.
    \item Stage 2: Close pairs showing strong signs of interaction.
    \item Stage 3: Well separated pairs with evidence of disturbance.
    \item Stage 4: Overlapping with morphological distortion.
\end{itemize}

% \vspace{-2mm}

\noindent Each stage is designed to capture a different part of interaction, where we expect different physical processes to be dominant.

In the first stage of interaction, the two galaxies remain separated with no morphological disturbance. We expect this to be because the two galaxies are approaching each other and about to begin their first pericentre pass. Therefore, we would not expect any of the usual effects of interaction to have started. This stage acts, essentially, as a control on our sample.

In stage 2, we observe morphology distortion as the two galaxies have crossed their first pericentre passage.

In stage 3, we expect to have captured those galaxies which have undergone their first passage of each other and be approaching their apocentre. There are then three future outcomes of this. First, the secondary galaxy is only just approaching its apocentre, it has just passed it's apocentre, or the galaxy is about to escape from the system. In any case, we observe severe morphological disturbance at this point, and would expect to begin seeing the effects of interaction on the galaxy. The severe torques exerted on the gas should lead to some initial effects of a starburst, and there could also be some ignition of nuclear activity.

Stage 4 expects to capture those galaxies which are actually undergoing coalescence or be at the point where they are about to. This is, from the literature, the point we would expect all the effects of interaction to be taking place.

The categorisation was conducted by the co-authors of this paper (names). Initially, 300 samples from the COSMOS sample are classified together, to reduce biases from the group and then split between the group evenly. This leads to 4,385 classifications being made by each co-author. To further quantify the bias and error involved in using different classifiers, a further 20 galaxies were classified as a group followed by each member of the group classifying individually the same 80 galaxies. The results of this is shown in Figure \ref{fig:std-dev-clsf}. It is important to note here, that the two stages with highest disagreement between classifiers is stages 1 and 4. \textcolor{red}{Explain why this could be?}

\section{Results \& Discussion}\label{results}
 - Show breakdown of stages. \\
 - What are their mass distributions? \\
 - Mass Ratios \\
 - Investigate SF with Mass \\
 - Investigate SF with Mass Ratio/Merger Type \\
 - Investigate SF with projected Separation \\
 - Control for Environment \\
 - Where do the interacting galaxies lie? \\
 - What does this all mean?

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{hyperref}
% \usepackage{enumitem}

@@

\section{Introduction}\label{introduction}
\noindent Galaxy interaction and merging is of fundamental importance to our current theories of galaxy formation and evolution. Our current cosmological theory, $\Lambda$CDM dictates that structure is expected to form and assemble hierarchically \citep{1978MNRAS.183..341W, 2013MNRAS.436.1765M}. Thus, understanding the underlying processes that occur in galaxy interaction and merging are paramount to understanding cosmological theories. It is well known that many underlying processes are triggered or enhanced due to galaxy interaction. The clearest result of interaction is the disruption of galactic morphology, the acceleration of disks to spheroidal systems and the formation of tidal features \citep{1972ApJ...178..623T, 1977ApJ...212..616T, 2005MNRAS.357..753G, 2009MNRAS.397..802H}, increases in star formation rates when compared to the general galaxy population \citep{1991ApJ...370L..65B, 2006ApJ...652...56B, 2014MNRAS.437.2137S, 2015ApJ...807L..16K} and, potentially, their rapid quenching \citep{2013MNRAS.430.1901H, 2023RAA....23i5026D}. There is also evidence to suggest that interaction plays a role in the activation of active galactic nuclei (AGN) \citep{2011MNRAS.418.2043E, 2015ApJ...806..219C, 2023MNRAS.523.4164H}. Therefore, with interaction having such an affect on the galaxies we see - over a long time period - understanding galaxy interaction is fundamental to understanding galaxy evolution

To better our understanding of galaxy interaction, it is often prudent to break them down into different stages of interaction. This relates to the dynamical time of the interaction, and the snapshot in the interaction we are observing. Previous works focused on simulations find that a sudden increase in star formation occurs early in the dynamical timescale of the interaction \citep{2008MNRAS.384..386C, 2019MNRAS.490.2139R}, however, confirming this definitive link observationally remains elusive \citep{2023arXiv230916531R}. The same studies could also be done for nuclear activation as well as quenching of interacting systems \citep{Papers on nuclear activation}. While we cannot yet directly measure when in the dynamical timescale star formation is enhanced from observations, we can approximate it by splitting different interactions into different stages. These stages relate directly to the dynamical timescale of the interaction. Often, they are split into four: gravitationally bounded pairs, approaching pericentre of the interaction, approaching apocentre of the interaction, and approaching coalescence.

To fully study the relationship between interaction stage and the underlying physical processes occurring we require large, statistically robust interacting galaxy samples. However, they are surprisingly difficult to create; with the first interacting galaxy catalogue (the \citet{1966ApJS...14....1A} catalogue) containing only 318 systems. More recent samples have been created either by machine learning classification \citep{2019A&A...626A..49P, 2023A&A...669A.141S}, visual classification by citizen scientists \citep{2010MNRAS.401.1043D} or by photometric parameterisation \citep{2004AJ....128..163L, 2023MNRAS.522....1N}. However, each sample has always been plagued by contamination and a loss of statistical significance when broken down into stages. It is notoriously difficult, from morphology alone, to select interacting or merging galaxies. Without redshift information it is difficult to discern if they are close together in 3D space or only by projection effects.

We elect to use the newly created interacting galaxy catalogue created by \citet{2023ApJ...948...40O} (hereafter, OR23) which found a sample of 21,926 interacting galaxy systems. This is the largest visually classified interacting galaxy catalogue to date. This sample was created using the new ESA Datalabs\footnote{ESA Datalabs: link} platform and the convolutional neural network machine 
\texttt{Zoobot} \citep{2022MNRAS.509.3966W, 2023JOSS....8.5312W}. This classified all sources in the Hubble Source Catalogue \citep{2016AJ....151..134W} from the Hubble Space Telescope (HST) archives into interacting or non-interacting. Once this was done, a rigorous process of de-duplication, contamination removal and visual inspection was conducted.

In this work, we cross match this catalogue with the Cosmic Evolutionary Survey (COSMOS) survey\footnote{DOI: \href{https://www.ipac.caltech.edu/doi/irsa/10.26131/IRSA178}{10.26131/IRSA178}}. This provides us with ancillary data for a subset of the OR23 catalogue. This ancillary data contains many galactic parameters of interest and we explore how they evolve with different stages of interaction. We define four distinct stages that occur over the course of the full dynamical timescale of interaction, from the two galaxies approaching each other to full coalescence. We primarily focus on the evolution of the stellar masses, star formation rates, surrounding environment and AGN classification from various COSMOS catalogues across the full dynamical timescale of interaction. With the photometric redshifts available from COSMOS, we also confirm a set of interacting galaxies and close pairs to investage further parameters of interest. These include the trend of projected separation of the galactic pairs on star formation rate as well as the change with mass ratio.

%, comparing to the work of \citet{2022ApJ...940....4S} a different study using COSMOS to investigate these latter relations.

This paper is laid out as follows in Section \ref{data} we explain the OR23 and COSMOS catalogues, and describe the process of catalogue matching especially focused on accurate de-duplication and cross-matching selection from the different catalogues.  Section \ref{staging} describes the methods by which we split the interacting systems into their different stages, and how we identify the secondary galaxies to each identified interacting galaxy. We also discuss our processes of calculating projected separations. In Section \ref{results} we describe and discuss our results, and their implications, and compare to previous works focused on investigating different stages of interaction. Finally, Section \ref{conclusion} we make concluding remarks and discuss future work to better our constraints.

@@

\section{DATA: Catalogue Matching \& Secondary Identification} \label{data}
\subsection{The O'Ryan+23 Catalogue}
\noindent We utilise the interacting galaxy catalogue created in OR23. While the full process of its creation is detailed there, we briefly describe it here. This sample of interacting galaxies was created using a convolutional neural network - \texttt{Zoobot} - to search all final product data in the \textit{Hubble} Legacy Archive. A regular convolutional neural network classifies images by extracting features from them, and using this to output a solution. \texttt{Zoobot} differs by extracting features from an image, and then reduces them to a generalised representation (essentially, a small feature vector). The power of this is that it makes \texttt{Zoobot} incredibly generalisable. This is further reinforced by training it on citizen scientist responses to the Galaxy Zoo: DECaLS \citep{2022MNRAS.509.3966W} project to every task in that project. Thus, it can answer a host of questions as well as be very easily finetuned to new datasets and classification tasks. In OR23, it was easily finetuned on \textit{Hubble} data by using classifications from all previous Galaxy Zoo projects \citep{2008MNRAS.389.1179L}.

Cutouts of each source were created as single-band \textit{F814W} images at 150 $\times$ 150 pixels across. This choice was motivated by it having the highest number of final product FITS files associated with it, while the image size was kept small due to the high number of cutouts required to be stored. Source coordinates were found in the \textit{Hubble} Source Catalogue \citep{2016AJ....151..134W}. This corresponded to just over 92 million cutouts to create and sources to classify. Due to issue of classifying interacting galaxies by morphology alone an extremely conservative prediction score from \textit{Zoobot} was required. This resulted in a very pure, but highly incomplete catalogue of 21,926 interacting galaxy systems. This is the largest catalogue of interacting galaxies directly found to date.

However, in the OR23 catalogue, only the coordinates and the \textit{Hubble} Source catalogue are provided. For ancillary data, this catalogue must be matched to different data repositories. This is not a trivial task, as a large proportion of the OR23 catalogue is known to not exist in any existing astrophysical data repository. In this work, we opt to match the OR23 catalogue to the COSMOS catalogues.

\subsection{The COSMOS2020 Catalogue}
We opt to match the OR23 catalogue with the catalogues of the COSMOS survey. Specifically, we elect to use the COSMOS2020 catalogue \citep{2022ApJS..258...11W}. This catalogue has a wealth of ancillary information for just over 1.7 million sources in a 2 square degree area of the sky. Each source found in this area has been analysed with well used astronomical software, for our purposes primarily LePhare \citep{1999MNRAS.310..540A, 2006A&A...457..841I} and EAzY \citep{2008ApJ...686.1503B}. These provide two different measurements of the parameters we wish to investigate and match with stage. The first is the stellar masses of each our cross matched galaxies and the estimated star formation rates. We opt to simply use one of these two measurements in our analysis. For the stellar mass we use the LePhare measurement and for the star formation rates we use the EAzY measurements. Using either the vice versa of this, both EAzY or both LePhare measurements led to the degeneracy's within the underlying model Spectral Energy Distribution (SED) templates being revealed. Therefore, using these two opposing parameters gave realistic approximations of these photometrically approximated values.

As stated previously, the OR23 catalogue only contained the source ID, right ascension and declination of each interacting galaxy. Therefore, we had to cross match these with the COSMOS2020 catalogue right ascension and declination measurements. To do this, we searched the COSMOS2020 catalogue for any sources within 10" of the OR23 sources. If more than one COSMOS2020 source was within this radius, we took the closest source to the OR23 coordinates as the correct source. Upon matching based on coordinates, we conducted a de-duplication of the resultant catalogue based on COSMOS2020 ID, and reassigned COSMOS2020 sources to each duplication. If no further COSMOS2020 sources were within 10" of the OR23 source, then this was considered not to be cross matched. We find that, after this cross matching and de-duplication, 3,786 of the OR23 sources exist in the COSMOS2020 catalogue.

We make two further cuts, based on the ancillary data available in the COSMOS2020 catalogue. First, there are many unphysical photometric measurements for the stellar mass or star formation rates. Therefore, we cut out these results as spurious. We only keep systems within a mass range of $6.5 \leq \log_{10} \text{M}_{*}(\text{M}_{\odot}) \leq 12.5$ and a star formation rate range of $-5 < \log_{10} \text{SFR} (M_{\odot}yr^{-1}) \leq 3.5$. Secondly, we institute a redshift cut of $z \leq 1.2$. This is to match the catalogue that will be later used to investigate the incidence of our sample of interacting galaxies with environment (detailed in Section \ref{data:environ}). Applying these cuts reduces our sample size to 3,689 interacting galaxies.

\subsection{Secondary Identification}
\noindent As stated previously, the OR23 catalogue only identified interacting galaxy systems, not individual interacting galaxies. Therefore, we must find the secondary in each interaction using the COSMOS2020 catalogue. The size of the cutouts in this catalogue were also small, and therefore, the secondary did not often appear in the cutout in that project. To find the secondaries, we apply three steps. First, a cutout surrounding the primary of each system was created. These cutouts were from the COSMOS cutout service, selecting HST-ACS tiles as the source of the cutouts in the $F814W$ filter. Each cutout had a 30" radius (corresponding to 1001 $\times$ 1001 pixels) centred and was centred on the primary galaxy. The original OR23 cutout was also displayed next to the enlarged cutout. We then searched the COSMOS2020 catalogue for all sources within the cutout. Their positions and redshifts were recorded and used to annotate the cutout.

We then visually assessed each annotated cutout, giving one of the four following classifications to each: system disturbed but secondary could not be identified; secondary could be identified; cannot confirm galaxy is interacting; null redshift; incorrect primary assigned. Our definition of a secondary galaxy was such that the secondary had to be within the cutout we were visually assessing and within $z=\pm0.1$ of the primary redshift. We use this cutoff as this is often used when calculating environment parameters \citep[e.g][]{Baldry 2006} or defining interacting galaxies by close pairs \citep[e.g][]{Shah et al. paper from 2022 / 2023} based on the uncertainties present in using photometric redshift estimates from SED fitting. We define a null redshift as one outwith our redshift limits or NaN. A minority of the cutouts we visually assessed were found to have the incorrect primary at the centre. This was a result of our cross matching based on radius with the COSMOS catalogue.

Using these definitions, we find that of the 3,689 original systems cross-matched with COSMOS2020 2,209 could not have their secondary identified, 834 had a clear secondary, 446 could not be reliably classified as an interacting galaxy, 248 had a null redshift and 149 were the incorrect primary. Figure \ref{fig:secondary_selection} shows an example of each of our classifications. Each secondary we identify was added to our sample, increasing our sample size to 3,829.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/cutouts_ex.pdf}
    \caption{An example of each visual classification made on the cross matched sample. From left to right these are the secondary could not be identified, the primary had a clear secondary, the primary could not be reliably classified as an interacting galaxy, the redshift was null and the incorrect primary was identified in OR+23. Based on these classifications, we either add the secondary galaxies to the sample or we remove the contamination from it. These images are 30" across using the COSMOS cutout service, selecting HST/ACS tiles as the basis for the observations in the $F814W$ filter.}
    \label{fig:secondary_selection}
\end{figure*}

% Define our classifications above.
While initially surprising that the majority of our systems could not have a secondary identified, we found that it was mostly due to limitations in the COSMOS catalogue or the way in which we had conducted our secondary identification. To define a secondary galaxy, the object had to have a ID associated with it in the COSMOS2020 catalogue. However, when two systems were very close together and small enough, often they would be identified under a single COSMOS ID despite being two separate systems. The same also occurred when two systems were merging or interacting. The tidal features connecting the systems would lead to only one of them being identified in the catalogue, or when two cores appeared in one merging system. Figure \ref{fig:secondary_selection} specifically an example of two systems being close enough together that they have been identified under a single COSMOS2020 entry. Figure \ref{fig:secondaries_found} shows this disparity with the different types of interaction we observe.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/visualisation_classification.pdf}
    \caption{Where a secondary could be identified at different stages in the interaction. The reasoning for such disparity in secondaries identified is due to the relative distance each the secondary would be from the primary at each stage. For a close pair, we often found the secondary galaxy, but a minority of these were so close together that the entire system was given a single COSMOS ID. The same was true for those interacting systems approaching pericentre or merging. When the secondary was near apocentre, often it would be outwith the cutout we were using for visual classification.}
    \label{fig:secondaries_found}
\end{figure}

Those galaxies which were found to be contamination (i.e., could not be reliably classified as interacting or having a null redshift) were removed from our sample. Galaxies that could not be reliably classified as interacting were those which showed little tidal distortion and had no systems nearby with a redshift approximately that of the primary. Often, these were systems that were overlapping but at different redshifts, systems with irregular morphologies of spiral arms or clumpy galaxies. There was also many systems that were at high redshift ($z \geq 1$) where the resolution of the cutouts meant that features could not be discerned visually.

Finally, we found that there were 149 systems which had been assigned to an incorrect primary. These were systems where the interacting galaxies were clearly in the cutout but some tidal debris or some nearby system had been assigned as the primary instead. This was a result of our catalogue matching based solely on separation between the OR23 coordinates and the COSMOS2020 source coordinates. We reassign these systems to the correct COSMOS IDs and then take them through the secondary identification process.

By the end of this pipeline, we find a sample size of 3,829 interacting galaxies. We conduct a de-duplication based on the COSMOS2020 ID, which reduces the sample back down to 3,547 interacting galaxies. This shows that many of the secondaries we had identified for the different systems were already found in the OR23 catalogue. The remaining systems are each visually confirmed interacting and disturbed galaxies based on their morphology and photometric redshift as measured in the COSMOS2020 catalogue.

\subsection{Finding Additional Systems}
\noindent As a result of using visual classification to find the secondary galaxy in each interaction, we were able to also confirm other interacting systems which had not been found in the OR23 catalogue. Primarily, these extra interacting galaxies were more systems that were involved in the interactions already recorded in OR23. The pipeline described previously was only used to find a secondary for each interacting galaxy, however in many cases more galaxies were involved. Other interacting galaxies that were added here were low redshift systems which would have appeared very large in the OR23 cutout. Details and tidal features were then lost in the original classification process. By looking at the larger COSMOS2020 cutouts used here, we were able to confirm these were interacting galaxies and add them to the catalogue. The final set of extra systems we added were often those that were close pairs, but physically in 3D space. While they had little tidal disruption - what the interacting galaxies were classified on in OR23 - with the photometric redshifts to hand, we could correctly identify them here.

In total, we found an extra 841 interacting systems that we could add to our sample. Upon conducting a de-duplication of these with the sample already found, this was reduced to 634 interacting systems. This gives us a total sample size of 4,181 which we use through the rest of this work.

\section{METHOD: Environment, AGN and Interaction Stage}
In this section, we describe how we separate our sample of 4,181 interacting galaxies into separate stages. Each stage is defined to capture a different part of the interaction, and a different part of the dynamical history of the interaction. We want to compare different stages in the interaction to different physical underlying processes, such as star formation rate. Such parameters can be found in the COSMOS2020 catalogue. However, two fundamental properties we also want to compare with is that of AGN fraction evolution with time and that of environment, and the impacts this might have.

The parameters required to calculate the AGN fraction are not found in the COSMOS2020, and we therefore must cross match with other catalogues to find the required parameters. We also describe the catalogue we use here to define the density parameter around each of our sources, and define how the environment is calculated there.

\subsection{Classifying Stage of Interaction}\label{sec:staging}
The primary goal of this work is to find if a relation exists between a host of galactic parameters, underlying physical processes and the stage of the interaction. We, therefore, need to define different stages for them. We have hinted at this in the previous section with Figure \ref{fig:secondaries_found}. We break our sample down into four stages:

\begin{itemize}
    \item Stage 1: Systems which are well-separated with no morphological disturbance.
    \item Stage 2: Close pairs showing either morphological distortion or physical connection by tidal features.
    \item Stage 3: Well separated pairs with morphological disturbance or isolated galaxies with clear tidal features.
    \item Stage 4: Highly disturbed systems with two or more cores within.
\end{itemize}

\noindent Examples of each stage can be found in Figure \ref{fig:stages}. Each stage is designed to capture a different part of the interaction history of a system in the total dynamical time.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/examples-stages.pdf}
    \caption{Examples of the four stages we split our interacting galaxy sample into. Stage 1: A close pair with confirmed redshift matching. Stage 2: Two distinct systems interacting with tidal features forming. Stage 3: A tidally disturbed system with no secondary present, likely at apocentre. Stage 4: A galaxy with multiple cores while highly disturbed. At the final stage before coalescence.}
    \label{fig:stages}
\end{figure}

Stage 1 of the interaction captures the first approach of the two systems. This is the time when the two galaxies potentials are beginning to have an effect on one another. They have no morphological disturbance yet and, in our sample, are most often those galaxies with distinct disks. Hence, this stage has the highest number of if identified secondaries with every system having a secondary that can be visually classified. In this sample, we would expect no effects from the interaction to have began yet.

Stage 2 of the interaction is defined as the point where the two galaxies in the interaction are approaching or just passing the pericentre of the tidal encounter. For those galaxies before pericentre, we would expect the beginning of morphological disturbance or some irregularity in the two systems. Once the galaxies have passed pericentre, we would define it by much higher morphological disturbance, distortion and tidal debris. This requirement for morphological disturbance is how we differentiate stage 2 from stage 1 interacting galaxies. Throughout this stage of the interaction, in our sample, we see significant overlap or connection of the two systems in the form of tidal features and morphological distortion.

Stage 3 interacting galaxies are described as those interacting systems where the systems have approached apocentre from each other. In many of cases of this stage, if the galaxies have sufficient velocity, the galaxies will escape each other and not coalesce. Hence, galaxies at this stage often could not have a secondary galaxy identified in our visual assessment. We purely describe a galaxy as stage 3 based on significant tidal and morphological disturbance. This stage has the highest potential for contamination by irregular galaxies or galaxies whose morphology has been disturbed by processes other than tidal interaction. We discuss the potential impacts of this contamination, and the limitations of using visual morphology as a proxy for stage in Section \ref{sec:limitations}.

Finally, stage 4 represents the final step of a galaxy interaction. If the two galaxies do not have sufficient velocity to escape one another, then they go on to coalesce and ultimately merge. We define this stage such that the sample is of galaxies just before this coalescence and the two cores remain distinct. However, it is important to note that this stage may be contaminated by some post-merger remnants. In this context, we define a stage 4 galaxy as one with significant morphological disturbance but with multiple cores present in the disks. At this stage, the interaction is at it's most violent with complete disruption of both galactic disks. We also expect star formation rate to increase significantly in this stage. This is important to note, as high star forming at specific redshifts in the $F814W$ filter may appear like a second core within a galactic disk. At this point, we would be observing a post-merger remnant rather than the point just before coalescence. Therefore, to mitigate potential contamination in from the post-merger stage we only classify stage 4 galaxies as those with highly disrupted disks and second cores close to the primary core (and not far out in the disk).

Our four stage approach is not a new one, and many other works have utilised it to differentiate different parts of the dynamical history of a galaxy interaction \citep{}. It is, however, important to note its limitations. First, this is a simplified view of the dynamical timescale of an interaction. We have assumed that an interaction begins, the pair have enough velocity to escape or they coalesce and that is representative of the full dynamical timescale. We have not considered systems where there are multiple flybys before coalescence or systems where there are multiple members at different stages of the interaction. This is a particular concern when looking at stages 2 and 3. For instance, the effects could be enhanced or diminished over multiple flybys, and we have no way to capture this.

However, with these simplifications in our assumptions, we will investigate the parameters which are proxies for the underlying physical processes and interpret them. We will investigate the whether we see have a bias in our systems based on environment, whether we find increased AGN activity and if we observe enhancement in star formation with stage. The star formation rate is already present in the COSMOS2020 catalogue from both EAzY and LePhare. For the environment parameter and whether AGN are detected, we cross match to two further catalogues. We describe these catalogues here, our process of cross matching and the new sample sizes we have in each subsample.

\subsection{Matching to Environment Catalogue}\label{data:environ}
\noindent There is no measure of the environmental density in the COSMOS2020 catalogue. Such a measure is often calculated in numerous ways, such as the N-nearest neighbour \citep{2006MNRAS.373..469B}, different Bayesian metrics \citep{2008ApJ...674L..13C} or estimating it from Voronoi Tesselation \citep{2021inas.book...57V}. However, in this work, we utilise the existing environmental density catalogue produced by \citet{2017ApJ...837...16D}. This catalogue was created specifically for the COSMOS survey, and has a measured density for all sources with mass $\log(\frac{M}{M_\odot}) \geq 9.6$ and out to a redshift of 1.2. In this catalogue, they calculate not only the density, but also the density parameter $\delta$ and assign each source to a field, filament or cluster classification. For a full description of how they calculate the environment and density field see \citet{2015ApJ...805..121D} and \citet{2017ApJ...837...16D}, but we will briefly describe it here.

To build the density field throughout the COSMOS field, they first construct a set of overlapping redshift slices. Within each slice, a subset of the galaxies are selected such that the median of the probability distribution function (PDF) of their photometric redshift is within it. Then, from this subset, they calculate the weighted surface density within the redshift slice. The weighting is based upon the PDF of the photmetric redshift present within the redshift slice. These weights significantly reduce the effects of projection effects. They then apply a weighted adaptive kernel smoothing using a 2D Gaussian kernel whose width changes based on the found local density of galaxies. Once this density field is created, the density around the sample galaxies can simply be interpolated across the density field based on the angular position and the redshift slice the sample galaxy is in (based on its photometric redshift PDF).
